---
title: "LacZ COSMIC Project"
author: "Matthew J. Meier"
subtitle: MutationalPatterns analysis - for response to reviewers
output:
  html_document:
    code_folding: show
    fig_caption: yes
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '4'
---

***
# Revision History
#### This version
Current version: 1  
Date report generated: `r format(Sys.time(), '%d %B, %Y')`  
Purpose of report: To test whether a different signature reconstruction algorithm applied to the lacZ COSMIC data produced robust signature compositions for each environmental mutagen in the study.  

#### Previous revisions

Previous version: N/A  
Revision requested by: N/A  
Justification for revision: N/A  

***
# Load libraries required for analysis
```{r setup,echo=TRUE,include=TRUE,message = FALSE}
library(MutationalPatterns)
library(GenomicRanges)
library(VariantAnnotation)
library(SomaticSignatures)
library(deconstructSigs)
library(ggplot2)
```

***
# Load data required for analysis
#### lacZ-corrected COSMIC signatures (V3) WITHOUT control
```{r load_v3_no_control}
lacZ_cosmic_signatures_v3_NO_control <- as.data.frame(t(read.table("data/lacZ_corrected_COSMIC_signatures_v3_without_control.txt",
                                                                sep="\t",
                                                                header=F,
                                                                stringsAsFactors = F)))
newSigs <- read.table("data/ColumnNamesforSignatures3withoutcontrol.txt", header=T)
rownames(lacZ_cosmic_signatures_v3_NO_control) <- colnames(newSigs)
colnames(lacZ_cosmic_signatures_v3_NO_control) <- colnames(signatures.cosmic)
```

#### lacZ-corrected COSMIC signatures (V3) with control
```{r load_v3}
lacZ_cosmic_signatures_v3_control <- as.data.frame(t(read.table("data/lacZ_corrected_COSMIC_signatures_v3_with_control.txt",
                                                     sep="\t",
                                                     header=T,
                                                     stringsAsFactors = F)))
colnames(lacZ_cosmic_signatures_v3_control) <- colnames(signatures.cosmic)
```

#### lacZ-corrected COSMIC signatures (V2) with control
```{r load_v2}
lacZ_cosmic_signatures_v2_control <- as.data.frame(t(read.table("data/lacZ_corrected_COSMIC_signatures_v2_with_control.txt",
                                                                     sep="\t",
                                                                     header=T,
                                                                     stringsAsFactors = F)))
colnames(lacZ_cosmic_signatures_v2_control) <- colnames(signatures.cosmic)
```

#### Mutation observations
```{r load_muts}
mutationData <- read.table("./mutation_data.txt",
                           header=T,
                           sep="\t")
mutationData$chr <- "lacZ"
colnames(mutationData)[2] <- "start"
mutationData$end <- mutationData$start

mutationDataVRange <- VRanges(seqnames = mutationData$chr,
                                ranges = IRanges(mutationData$start,
                                end = mutationData$end),
                                ref = mutationData$Ref,
                                alt = mutationData$Alt,
                                sampleNames = mutationData$Group)
# Sanity check
head(mutationData)
```
***
# Data processing
#### Generate motif matrix for mutation data
```{r motif_matrix}
lacZ.fa = FaFile(paste("data/lacZ.fa", sep=""))
mutationRanges <- mutationContext(mutationDataVRange, lacZ.fa)
mut_matrix = as.matrix(motifMatrix(mutationRanges))
```

#### Do fitting of signatures
```{r fitting}
fit_res_v2 <- fit_to_signatures(mut_matrix, t(lacZ_cosmic_signatures_v2_control))
fit_res_v3 <- fit_to_signatures(mut_matrix, t(lacZ_cosmic_signatures_v3_control))
fit_res_v3_no_control <- fit_to_signatures(mut_matrix, t(lacZ_cosmic_signatures_v3_NO_control))
```

#### Cosine similarity calculation for reconstructions
V2 COSMIC Signatures
```{r cosine_similarities_V2}
cos_sim_v2 <- cos_sim_matrix(fit_res_v2$reconstructed, mut_matrix)
cos_sim_v2_lacZ <- cos_sim_matrix(fit_res_v2$reconstructed, t(lacZ_cosmic_signatures_v2_control))

```
V3 COSMIC Signatures
```{r cosine_similarities_V3}
cos_sim_v3 <- cos_sim_matrix(fit_res_v3$reconstructed, mut_matrix)
cos_sim_v3_lacZ <- cos_sim_matrix(fit_res_v3$reconstructed, t(lacZ_cosmic_signatures_v3_control))

```

V3 COSMIC Signatures, no control signature
```{r cosine_similarities_V3_no_control}
cos_sim_v3_no_control <- cos_sim_matrix(fit_res_v3_no_control$reconstructed, mut_matrix)
cos_sim_v3_no_control_lacZ <- cos_sim_matrix(fit_res_v3$reconstructed, t(lacZ_cosmic_signatures_v3_NO_control))
```

***

# Plots
```{r cutoff, eval=TRUE, echo=FALSE}
cutoff=0.0
```

Using a cutoff value of `r cutoff`  

## Contributions and cosine similarities
There are four plots for each version of the COSMIC signatures.  
  
The first two plots show relative contribution of each signature (first in bar chart, then as a heat map).  
  
The final two plots show heatmaps of cosine similarity: first for the reconstructed signatures vs. actual mutations, then for the reconstructed signatures vs. the lacZ-COSMIC signatures used for the reconstruction.  
#### V2 Signatures, with control signature
```{r plot_V2, fig.cap="Contribution of each signature (V2 COSMIC)", fig.width=10}
selectv2 <- which(rowSums(fit_res_v2$contribution) > cutoff)
plot_contribution(fit_res_v2$contribution[selectv2,],
                  lacZ_cosmic_signatures_v2_control[,selectv2],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v2$contribution,
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V2_cos1, fig.cap="Cosine similarity: reconstructed signatures vs. mutations", fig.width=10}
plot_cosine_heatmap(cos_sim_v2)
```

```{r plot_V2_cos2, fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v2_lacZ)
```

#### V3 Signatures, with control signature
```{r plot_V3, fig.cap="Contribution of each signature (V3 COSMIC)", fig.width=10}
selectv3 <- which(rowSums(fit_res_v3$contribution) > cutoff)
theme_set(theme(legend.position="bottom"))
plot_contribution(fit_res_v3$contribution[selectv3,],
                  lacZ_cosmic_signatures_v3_control[,selectv3],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v3$contribution,
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V3_cos1, fig.cap="Cosine similarity: reconstructed signatures vs. mutations", fig.width=10}
plot_cosine_heatmap(cos_sim_v3)
```

```{r plot_V3_cos2, fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_lacZ)
```

#### V3 Signatures, without control signature
```{r plot_v3_no_control, fig.cap="Contribution of each signature (V3 COSMIC, no control)", fig.width=10}
selectv3_no_control <- which(rowSums(fit_res_v3_no_control$contribution) > cutoff)
plot_contribution(fit_res_v3_no_control$contribution[selectv3_no_control,],
                  lacZ_cosmic_signatures_v3_NO_control[,selectv3_no_control],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v3_no_control$contribution,
                          cluster_samples = TRUE,
                          method = "complete") + theme(legend.position = "bottom")
```

```{r plot_V3_no_control_cos1, fig.cap="Cosine similarity: reconstructed signatures vs. mutations", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_no_control)
```

```{r plot_V3_no_control_cos2, fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_no_control_lacZ)
```

## Reconstructed signatures
The following plots show the top panel as original mutation data, the middle panel as the computationally reconstructed signature, and the bottom panel as the difference. Note that the y-axis is an order of magnitude smaller for the bottom panel.
#### V2 COSMIC
```{r plot_reconstructions_V2}
for (i in 1:ncol(mut_matrix)) {
print(plot_compare_profiles(mut_matrix[,i], fit_res_v2$reconstructed[,i],
                      profile_names = c(colnames(mut_matrix)[i], "Reconstructed"),
                      condensed = TRUE))
}
```

#### V3 COSMIC
```{r plot_reconstructions_V3}
for (i in 1:ncol(mut_matrix)) {
print(plot_compare_profiles(mut_matrix[,i], fit_res_v3$reconstructed[,i],
                      profile_names = c(colnames(mut_matrix)[i], "Reconstructed"),
                      condensed = TRUE))
}
```

#### V3 COSMIC, no control signature
```{r plot_reconstructions_V3_no_control}
for (i in 1:ncol(mut_matrix)) {
print(plot_compare_profiles(mut_matrix[,i], fit_res_v3_no_control$reconstructed[,i],
                      profile_names = c(colnames(mut_matrix)[i], "Reconstructed"),
                      condensed = TRUE))
}
```
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
