---
title: "LacZ COSMIC Project"
author: "Matthew J. Meier"
subtitle: MutationalPatterns analysis - for response to reviewers
output:
  html_document:
    code_folding: show
    fig_caption: yes
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: no
  pdf_document:
    documentclass: hitec
    toc: yes
    toc_depth: '4'
    pandoc_args: --listings
    includes:
      in_header: preamble.tex
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), paste0("lacZ_COSMIC_MutationalPatterns_",format(Sys.time(), '%d-%m-%Y.%H.%M'),".html"))) })
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)
```

***
# Revision History
#### This version
Previous version: 4   
Date report generated: `r format(Sys.time(), '%d %B, %Y')`  
Revision requested by: Francesco Marchetti  
Justification(s) for revision:

* Downsample data to illustrate robustness

#### Previous revisions

Previous version: 3   
Date report generated: 16 June 2020  
Revision requested by: Francesco Marchetti  
Justification(s) for revision:

* Resample data to display random reassignment of the 96 triplet mutations within each sample  
* Show relative contributions with and without Sanger data (i.e., downsampling)  

Previous version: 2  
Date report generated: 27 May, 2020  
Revision requested by: Francesco Marchetti  
Justification(s) for revision:

* Show a table of cosine similarities (between lacZ COSMIC signatures and the reconstructed signatures for each mutagen) above the threshold of 0.5 as done in previous analysis
* Combine NGS and Sanger data

Additional changes:

* Removed instances where the reference base in mutation data had a mismatch from the *lacZ* reference
* Removed the display of signatures in figures that had zero contribution to a particular mutational catalog
* Show tabular output for relative contribution for reconstruction

Previous version: 1  
Date of previous version: 26 May, 2020  
Purpose of report: To test whether a different signature reconstruction algorithm applied to the lacZ COSMIC data produced robust signature compositions for each environmental mutagen in the study.  


***
# Load libraries required for analysis
```{r setup,echo=TRUE,include=TRUE,message = FALSE}
library(MutationalPatterns)
library(GenomicRanges)
library(VariantAnnotation)
library(SomaticSignatures)
library(deconstructSigs)
library(ggplot2)
library(knitr)
library(dplyr)
library(kableExtra)
```

***
# Load data required for analysis
#### lacZ-corrected COSMIC signatures (V3) WITHOUT control
```{r load_v3_no_control}
lacZ_cosmic_signatures_v3_NO_control <- as.data.frame(t(read.table("data/lacZ_corrected_COSMIC_signatures_v3_without_control.txt",
                                                                sep="\t",
                                                                header=F,
                                                                stringsAsFactors = F)))
newSigs <- read.table("data/ColumnNamesforSignatures3withoutcontrol.txt", header=T)
rownames(lacZ_cosmic_signatures_v3_NO_control) <- colnames(newSigs)
colnames(lacZ_cosmic_signatures_v3_NO_control) <- colnames(signatures.cosmic)
```

#### lacZ-corrected COSMIC signatures (V3) with control
```{r load_v3}
lacZ_cosmic_signatures_v3_control <- as.data.frame(t(read.table("data/lacZ_corrected_COSMIC_signatures_v3_with_control.txt",
                                                     sep="\t",
                                                     header=T,
                                                     stringsAsFactors = F)))
colnames(lacZ_cosmic_signatures_v3_control) <- colnames(signatures.cosmic)
```

#### lacZ-corrected COSMIC signatures (V2) with control
```{r load_v2}
lacZ_cosmic_signatures_v2_control <- as.data.frame(t(read.table("data/lacZ_corrected_COSMIC_signatures_v2_with_control.txt",
                                                                     sep="\t",
                                                                     header=T,
                                                                     stringsAsFactors = F)))
colnames(lacZ_cosmic_signatures_v2_control) <- colnames(signatures.cosmic)
```

#### Mutation observations
```{r load_muts}
mutationData <- read.table("./mutation_data.txt",
                           header=T,
                           sep="\t")
mutationData$chr <- "lacZ"
colnames(mutationData)[2] <- "start"
mutationData$end <- mutationData$start
# Changed in version 2 to use chemical as sample names instead (i.e., not split by NGS vs Sanger)
mutationData$chemical <- gsub("_.*$", "", mutationData$Group)
mutationDataVRange <- VRanges(seqnames = mutationData$chr,
                                ranges = IRanges(mutationData$start,
                                end = mutationData$end),
                                ref = mutationData$Ref,
                                alt = mutationData$Alt,
                                sampleNames = mutationData$chemical,)
mutationDataVRangeSanger <- VRanges(seqnames = mutationData$chr,
                                ranges = IRanges(mutationData$start,
                                end = mutationData$end),
                                ref = mutationData$Ref,
                                alt = mutationData$Alt,
                                sampleNames = mutationData$Group,)
# Sanity check
knitr::kable(head(mutationData),
             caption="Sanity check for mutation data") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

***
# Data processing
#### Generate motif matrix for mutation data
Added in version 2: check mutation data reference against our lacZ reference sequence.  
```{r motif_matrix}
lacZ.fa = FaFile(paste("data/lacZ.fa", sep=""))
refbase <- getSeq(lacZ.fa)
flagpos <- vector()
for (i in 1:length(start(mutationDataVRange))) {
  if (ref(mutationDataVRange)[i] != as.character(refbase$lacZ[start(mutationDataVRange)[i]])) {
    flagpos=c(flagpos, i)
    }
}
# Remove positions that don't match the reference
knitr::kable(mutationDataVRange[flagpos],
             caption = "The following mutation calls will be removed because the reference sequence of lacZ does not match the reference from the mutation data.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
mutationDataVRange[flagpos] <- NULL
mutationDataVRangeSanger[flagpos] <- NULL
mutationRanges <- mutationContext(mutationDataVRange, lacZ.fa)
mutationRangesSanger <- mutationContext(mutationDataVRangeSanger, lacZ.fa)
set.seed(42)
mutationRangesDownsampled <- sample(mutationRanges, round(length(mutationRanges)/2), replace=F)
mut_matrix = as.matrix(motifMatrix(mutationRanges, normalize=FALSE))
mut_matrix_norm = as.matrix(motifMatrix(mutationRanges, normalize=TRUE))
mut_matrixSanger = as.matrix(motifMatrix(mutationRangesSanger, normalize=FALSE))
mut_matrixDownsampled = as.matrix(motifMatrix(mutationRangesDownsampled, normalize=FALSE))
mut_df = data.frame(t(motifMatrix(mutationRanges, normalize=TRUE)))
mut_dfSanger = data.frame(t(motifMatrix(mutationRangesSanger, normalize=TRUE)))
mut_dfDownsampled = data.frame(t(motifMatrix(mutationRangesDownsampled, normalize=TRUE)))
colnames(mut_df) <- colnames(signatures.cosmic)
colnames(mut_dfSanger) <- colnames(signatures.cosmic)
```

#### Do fitting of signatures with MutationalPatterns

```{r fitting}
originalRownames <- rownames(mut_matrix)
set.seed(42)
mut_matrix_resampled <- mut_matrix[sample(nrow(mut_matrix)),]
row.names(mut_matrix_resampled) <- originalRownames
fit_res_v2 <- fit_to_signatures(mut_matrix, t(lacZ_cosmic_signatures_v2_control))
fit_res_v3 <- fit_to_signatures(mut_matrix, t(lacZ_cosmic_signatures_v3_control))
fit_res_v3_resampled <- fit_to_signatures(mut_matrix_resampled, t(lacZ_cosmic_signatures_v3_control))
fit_res_v3_downsampled <- fit_to_signatures(mut_matrixDownsampled, t(lacZ_cosmic_signatures_v3_control))
fit_res_v3Sanger <- fit_to_signatures(mut_matrixSanger, t(lacZ_cosmic_signatures_v3_control))
fit_res_v3_no_control <- fit_to_signatures(mut_matrix, t(lacZ_cosmic_signatures_v3_NO_control))
```

#### Do fitting of signatures with deconstructSigs

```{r fitting_deconstructSigs}
deconstructSigsResults <- list()

for (i in 1:nrow(mut_df)) {
	deconstructSigsResults[[i]] <- whichSignatures(tumor.ref=mut_df[i,],
                                              signatures.ref=lacZ_cosmic_signatures_v3_control,
                                              sample.id=rownames(mut_df[i,]))
}

dsWeights <- matrix(ncol = nrow(lacZ_cosmic_signatures_v3_control), nrow = nrow(mut_df))
dsProduct <- matrix(ncol = ncol(lacZ_cosmic_signatures_v3_control), nrow = nrow(mut_df))
vcfNames <- vector()

for (i in 1:nrow(mut_df)) {
	dsWeights[i,] <- as.numeric(deconstructSigsResults[[i]]$weights)
	dsProduct[i,] <- as.numeric(deconstructSigsResults[[i]]$product)
 	vcfNames <- c(vcfNames, rownames(deconstructSigsResults[[i]]$weights))
}

rownames(dsWeights) <- vcfNames
colnames(dsWeights) <- names(deconstructSigsResults[[1]]$weights)
rownames(dsProduct) <- vcfNames
colnames(dsProduct) <- colnames(deconstructSigsResults[[1]]$product)

```

#### Do fitting of signatures with deconstructSigs including Sanger data

```{r fitting_deconstructSigsSanger}
deconstructSigsResultsSanger <- list()

for (i in 1:nrow(mut_dfSanger)) {
	deconstructSigsResultsSanger[[i]] <- whichSignatures(tumor.ref=mut_dfSanger[i,],
                                              signatures.ref=lacZ_cosmic_signatures_v3_control,
                                              sample.id=rownames(mut_dfSanger[i,]))
}

dsWeightsSanger <- matrix(ncol = nrow(lacZ_cosmic_signatures_v3_control), nrow = nrow(mut_dfSanger))
dsProductSanger <- matrix(ncol = ncol(lacZ_cosmic_signatures_v3_control), nrow = nrow(mut_dfSanger))
vcfNamesSanger <- vector()

for (i in 1:nrow(mut_dfSanger)) {
	dsWeightsSanger[i,] <- as.numeric(deconstructSigsResultsSanger[[i]]$weights)
	dsProductSanger[i,] <- as.numeric(deconstructSigsResultsSanger[[i]]$product)
 	vcfNamesSanger <- c(vcfNamesSanger, rownames(deconstructSigsResultsSanger[[i]]$weights))
}

rownames(dsWeightsSanger) <- vcfNamesSanger
colnames(dsWeightsSanger) <- names(deconstructSigsResultsSanger[[1]]$weights)
rownames(dsProductSanger) <- vcfNamesSanger
colnames(dsProductSanger) <- colnames(deconstructSigsResultsSanger[[1]]$product)

```

#### Table of relative signature contributions
```{r tables}
# Calculate relative contributions
df <- fit_res_v3$contribution
df_downsampled <- fit_res_v3_downsampled$contribution
df_relative <- apply(df, 2, function(x){x/sum(x)})
df_relative_downsampled <- apply(df_downsampled, 2, function(x){x/sum(x)})

difference <- df_relative - df_relative_downsampled
mean(abs(difference[difference > 0]))

difference_for_plotting <- t(difference)
difference_for_plotting <- data.frame(difference_for_plotting, chemical=row.names(difference_for_plotting))
difference_for_plotting <- tidyr::pivot_longer(as.data.frame(difference_for_plotting),
                                               cols=-chemical,
                                               names_to = "signature",
                                               values_to = "proportion")
str(difference_for_plotting)
knitr::kable(df_relative,
             caption = "Relative contribution of each signature to the mutational catalog for each mutagen",
             digits = 2,
             fixed_thead = T)  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "720px")
```

#### Cosine similarity calculation for reconstructions
V2 COSMIC Signatures
```{r cosine_similarities_V2}
cos_sim_v2 <- cos_sim_matrix(fit_res_v2$reconstructed, mut_matrix)
cos_sim_v2_lacZ <- cos_sim_matrix(fit_res_v2$reconstructed, t(lacZ_cosmic_signatures_v2_control))

```
  
V3 COSMIC Signatures, no control signature
```{r cosine_similarities_V3_no_control}
cos_sim_v3_no_control <- cos_sim_matrix(fit_res_v3_no_control$reconstructed, mut_matrix)
cos_sim_v3_no_control_lacZ <- cos_sim_matrix(fit_res_v3$reconstructed, t(lacZ_cosmic_signatures_v3_NO_control))
```
  
V3 COSMIC Signatures  
  
Modified for version 2 of this analysis to add cutoff of 0.5
  
```{r cosine_similarities_V3}
cos_sim_v3 <- cos_sim_matrix(fit_res_v3$reconstructed, mut_matrix)
cos_sim_v3_lacZ <- cos_sim_matrix(fit_res_v3$reconstructed, t(lacZ_cosmic_signatures_v3_control))

cos_sim_v3S <- cos_sim_matrix(fit_res_v3Sanger$reconstructed, mut_matrixSanger)
cos_sim_v3_lacZS <- cos_sim_matrix(fit_res_v3Sanger$reconstructed, t(lacZ_cosmic_signatures_v3_control))

cos_sim_v3_resampled <- cos_sim_matrix(fit_res_v3_resampled$reconstructed, mut_matrix)
cos_sim_v3_lacZ_resampled <- cos_sim_matrix(fit_res_v3_resampled$reconstructed, t(lacZ_cosmic_signatures_v3_control))

cos_sim_v3_downsampled <- cos_sim_matrix(fit_res_v3_downsampled$reconstructed, mut_matrixDownsampled)
cos_sim_v3_lacZ_downsampled <- cos_sim_matrix(fit_res_v3_downsampled$reconstructed, t(lacZ_cosmic_signatures_v3_control))


cos_sim_v3_lacZ_filtered <- cos_sim_v3_lacZ
cos_sim_v3_lacZ_filtered[cos_sim_v3_lacZ_filtered<0.5] <- NA
```

V3 COSMIC Signatures: deconstructSigs (version 3)
```{r cosine_similarities_deconstructSigs}
cos_sim_v3_ds <- cos_sim_matrix(t(dsProduct), mut_matrix)
cos_sim_v3_lacZ_ds <- cos_sim_matrix(t(dsProduct), t(lacZ_cosmic_signatures_v3_control))

cos_sim_v3_lacZ_ds_filtered <- cos_sim_v3_lacZ_ds
cos_sim_v3_lacZ_ds_filtered[cos_sim_v3_lacZ_ds_filtered<0.5] <- NA
```

```{r cosine_similarities_deconstructSigs_Sanger}
cos_sim_v3_dsSanger <- cos_sim_matrix(t(dsProductSanger), mut_matrix)
cos_sim_v3_lacZ_dsSanger <- cos_sim_matrix(t(dsProductSanger), t(lacZ_cosmic_signatures_v3_control))

cos_sim_v3_lacZ_dsSanger_filtered <- cos_sim_v3_lacZ_dsSanger
cos_sim_v3_lacZ_dsSanger_filtered[cos_sim_v3_lacZ_dsSanger_filtered<0.5] <- NA
```

#### Tables showing cosine similarities

```{r cosine_similarities_V3_table1, tidy=TRUE}
# Shown as rounded to two decimals
knitr::kable(t(cos_sim_v3_lacZ),
             caption = "Cosine similarities between reconstructed signatures and lacZ COSMIC signatures",
             digits = 2,
             fixed_thead = T)  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "720px")
```

```{r cosine_similarities_table_rounded, tidy=TRUE}
cos_sim_v3_lacZ_filtered_rounded <- round(cos_sim_v3_lacZ_filtered,3)
knitr::kable(t(cos_sim_v3_lacZ_filtered_rounded),
             caption = "Cosine similarities between reconstructed signatures and lacZ COSMIC signatures, filtered at cosine similarity of 0.5",
             digits = 2,
             fixed_thead = T) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "720px")
  
```
  

***

# Plots
```{r cutoff, eval=TRUE, echo=FALSE}
cutoff=0.001
```

Using a cutoff value of `r cutoff`  

## Contributions and cosine similarities
There are four plots for each version of the COSMIC signatures.  
  
The first two plots show relative contribution of each signature (first in bar chart, then as a heat map).  
  
The final two plots show heatmaps of cosine similarity: first for the reconstructed signatures vs. actual mutations, then for the reconstructed signatures vs. the lacZ-COSMIC signatures used for the reconstruction.  

#### V2 Signatures, with control signature
```{r plot_V2, fig.align="center", fig.cap="Contribution of each signature (V2 COSMIC)", fig.width=10}
selectv2 <- which(rowSums(fit_res_v2$contribution) > cutoff)
plot_contribution(fit_res_v2$contribution[selectv2,],
                  lacZ_cosmic_signatures_v2_control[selectv2,],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v2$contribution[selectv2,],
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V2_cos1, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. mutations", fig.width=10}
plot_cosine_heatmap(cos_sim_v2)
```

```{r plot_V2_cos2, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v2_lacZ)
```

#### V3 Signatures, with control signature, Sanger data combined
```{r plot_V3, fig.align="center", fig.width=10}
selectv3 <- which(rowSums(fit_res_v3$contribution) > cutoff)
theme_set(theme(legend.position="bottom"))
plot_contribution(fit_res_v3$contribution[selectv3,],
                  lacZ_cosmic_signatures_v3_control[selectv3,],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v3$contribution[selectv3,],
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V3_cos1, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. mutations", fig.width=10}
plot_cosine_heatmap(cos_sim_v3)
```

```{r plot_V3_cos2, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_lacZ)
```

#### V3 Signatures, with control signature, Sanger data analyzed separately
```{r plot_V3_sangerdata, fig.align="center", fig.width=10}
selectv3S <- which(rowSums(fit_res_v3Sanger$contribution) > cutoff)
theme_set(theme(legend.position="bottom"))
plot_contribution(fit_res_v3Sanger$contribution[selectv3S,],
                  lacZ_cosmic_signatures_v3_control[selectv3S,],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v3Sanger$contribution[selectv3S,],
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V3_cos1_sangerdata, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. mutations", fig.width=10}
plot_cosine_heatmap(cos_sim_v3S)
```

```{r plot_V3_cos2_sangerdata, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_lacZS)
```

#### V3 Signatures, with control signature, analyzed by deconstructSigs whichSignatures function, Sanger data combined
```{r plot_V3_deconstructSigs, fig.align="center", fig.width=10}
selectv3DS <- which(rowSums(t(dsWeights)) > cutoff)
theme_set(theme(legend.position="bottom"))
plot_contribution(t(dsWeights)[selectv3DS,],
                  lacZ_cosmic_signatures_v3_control[selectv3DS,],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(t(dsWeights)[selectv3DS,],
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V3_cos1_deconstructSigs, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. mutations", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_ds)
```

```{r plot_V3_cos2_deconstructSigs, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_lacZ_ds)
```

#### V3 Signatures, with control signature, analyzed by deconstructSigs whichSignatures function, Sanger data analyzed separately
```{r plot_V3_sangerdata_deconstructSigs, fig.align="center", fig.width=10}
selectv3DSS <- which(rowSums(t(dsWeightsSanger)) > cutoff)
theme_set(theme(legend.position="bottom"))
plot_contribution(t(dsWeightsSanger)[selectv3DSS,],
                  lacZ_cosmic_signatures_v3_control[selectv3DSS,],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(t(dsWeightsSanger)[selectv3DSS,],
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V3_cos1_sangerdata_deconstructSigs, fig.align="center", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_dsSanger)
```

```{r plot_V3_cos2_sangerdata_deconstructSigs, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_lacZ_dsSanger)
```


#### V3 Signatures, with control signature, randomized (shuffled) data
```{r plot_V3_resampled, fig.align="center", fig.width=10}
selectv3resampled <- which(rowSums(fit_res_v3_resampled$contribution) > cutoff)
theme_set(theme(legend.position="bottom"))
plot_contribution(fit_res_v3_resampled$contribution[selectv3resampled,],
                  lacZ_cosmic_signatures_v3_control[selectv3resampled,],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v3_resampled$contribution[selectv3resampled,],
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V3_cos1_resampled, fig.align="center", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_resampled)
```

```{r plot_V3_cos2_resampled, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_lacZ_resampled)
```

#### V3 Signatures, with control signature, downsampled data
```{r plot_V3_downsampled, fig.align="center", fig.width=10}
selectv3downsampled <- which(rowSums(fit_res_v3_downsampled$contribution) > cutoff)
theme_set(theme(legend.position="bottom"))
plot_contribution(fit_res_v3_downsampled$contribution[selectv3downsampled,],
                  lacZ_cosmic_signatures_v3_control[selectv3downsampled,],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v3_downsampled$contribution[selectv3downsampled,],
                          cluster_samples = TRUE,
                          method = "complete")
```

```{r plot_V3_cos1_downsampled, fig.align="center", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_downsampled)
```

```{r plot_V3_cos2_downsampled, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_lacZ_downsampled)
```

The downsampled data has some differences in relative proportion - here is their histogram:

```{r plot_difference_downsampled}

ggplot(difference_for_plotting[difference_for_plotting$proportion > 0,], aes(x=abs(proportion))) + 
  geom_histogram(bins=10) +
  facet_wrap(~chemical)

ggplot(difference_for_plotting[difference_for_plotting$proportion > 0,], aes(x=abs(proportion))) + 
  geom_histogram(bins=10) +
  facet_wrap(~signature)

ggplot(difference_for_plotting[difference_for_plotting$proportion !=0,], aes(x = proportion,
                                                                             y = rev(signature),
                                                                             color=chemical)) +
          geom_point(position="jitter") +
          theme_bw() +
          scale_y_discrete(element_blank()) +
          scale_x_continuous(name = "% change in relative contribution") +
          theme(text = element_text(size=15))

ggplot(difference_for_plotting[difference_for_plotting$proportion !=0,], aes(x = chemical,
                                                                             y = proportion*100)) +
          geom_jitter(width=0.1, alpha=0.5) +
          theme_bw() +
          scale_y_continuous(name = "% change in relative contribution",
                             breaks = seq(-50, 50, by=10)) +
          theme(text = element_text(size=15),
                legend.position = "none",
                axis.text.x = element_blank()) +
  scale_x_discrete(name="Mutagen") +
  facet_grid(~chemical, scales="free_x")


```

#### V3 Signatures, without control signature

```{r plot_v3_no_control, fig.align="center", fig.cap="Contribution of each signature (V3 COSMIC, no control)", fig.width=10}
selectv3_no_control <- which(rowSums(fit_res_v3_no_control$contribution) > cutoff)
plot_contribution(fit_res_v3_no_control$contribution[selectv3_no_control,],
                  lacZ_cosmic_signatures_v3_NO_control[selectv3_no_control,],
                  coord_flip = TRUE,
                  mode = "relative") +
  theme(legend.position="bottom")
plot_contribution_heatmap(fit_res_v3_no_control$contribution[selectv3_no_control,],
                          cluster_samples = TRUE,
                          method = "complete") + theme(legend.position = "bottom")
```

```{r plot_V3_no_control_cos1, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. mutations", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_no_control)
```

```{r plot_V3_no_control_cos2, fig.align="center", fig.cap="Cosine similarity: reconstructed signatures vs. lacZ signatures", fig.width=10}
plot_cosine_heatmap(cos_sim_v3_no_control_lacZ)
```

## Reconstructed signatures
The following plots show the top panel as original mutation data, the middle panel as the computationally reconstructed signature, and the bottom panel as the difference. Note that the y-axis is an order of magnitude smaller for the bottom panel.
#### V2 COSMIC
```{r plot_reconstructions_V2}
for (i in 1:ncol(mut_matrix)) {
print(plot_compare_profiles(mut_matrix[,i], fit_res_v2$reconstructed[,i],
                      profile_names = c(colnames(mut_matrix)[i], "Reconstructed"),
                      condensed = TRUE))
}
```

#### V3 COSMIC
```{r plot_reconstructions_V3}
for (i in 1:ncol(mut_matrix)) {
print(plot_compare_profiles(mut_matrix[,i], fit_res_v3$reconstructed[,i],
                      profile_names = c(colnames(mut_matrix)[i], "Reconstructed"),
                      condensed = TRUE))
}
```

#### V3 COSMIC, no control signature
```{r plot_reconstructions_V3_no_control}
for (i in 1:ncol(mut_matrix)) {
print(plot_compare_profiles(mut_matrix[,i], fit_res_v3_no_control$reconstructed[,i],
                      profile_names = c(colnames(mut_matrix)[i], "Reconstructed"),
                      condensed = TRUE))
}
```

# Session Info
```{r sessionInfo, echo=FALSE, tidy=TRUE}
sessionInfo()
```

<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
